Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-07 paper.tex
first additions to section 3

Note that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of expressions. A verification condition is a logical formula (constructed from the bounded program and desired correctness properties) whose validity implies that the program's behaviour agrees with its specification~\cite{Bradley07}. Correctness properties in programs can be specified by the user via \textit{assert} statements or automatically generated from a specification language~\cite{Thomas01}. If all of a bounded program's VCs are valid, then a program is in compliance with its specification up to the given bound.

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). One may notice that the SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

Apart from that, this program has an exponential number of paths, since each element of $x$ can be either greater than one or less than or equal to one. Despite the large number of paths through that program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$, such that $\psi$ is satisfiable if and only if the assertion $a<=N$ fails. One may also notice that BMC still encodes program stages with a size that grows linearly with $N$. More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic, as follows~\cite{Bradley07}:
%
\begin{equation}              
\label{verification-condition-example}
\psi := \left [ \begin{array}{ll} 
                a_{1} = N  \\
                \wedge \, \, a_{2} = ite\left(x[0] > 1, a_{1} - 1, a_{1}\right) \\ 
                \wedge \, \, a_{3} = ite\left(x[1] > 1, a_{2} - 1, a_{2}\right) \\
                \wedge \, \, a_{4} = ite\left(x[2] > 1, a_{3} - 1, a_{3}\right) \\
                \wedge \ldots \\
                \wedge \, \, a_{N+1} = ite\left(x[N-1] > 1, a_{N} - 1 , a_{N}\right) \\
                \wedge \neg \left(a_{N+1} \leq N\right) \\
              \end{array} \right ]  \\
              \\           
\end{equation}

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).

>>>>

One may notice that BMC analyzes only bounded program runs, but generates verification conditions (VCs) that reflect the exact path in which a statement is executed, the context in which a given function is called, and the bit-accurate representation of expressions. A verification condition is a logical formula (constructed from the bounded program and desired correctness properties) whose validity implies that the program's behaviour agrees with its specification~\cite{Bradley07}. Correctness properties in programs can be specified by the user via \textit{assert} statements or automatically generated from a specification language~\cite{Thomas01}. If all of a bounded program's VCs are valid, then a program is in compliance with its specification up to the given bound.

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). The SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

Apart from that, this program has an exponential number of paths, since each element of $x$ can be either greater than one or less than or equal to one. Despite the large number of paths through that program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$, such that $\psi$ is satisfiable if and only if the assertion $a<=N$ fails. One may also notice that BMC still encodes program stages with a size that grows linearly with $N$. More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into $\psi$ using a decidable fragment of first-order logic, as follows~\cite{Bradley07}:
%
\begin{equation}              
\label{verification-condition-example}
\psi := \left [ \begin{array}{ll} 
                a_{1} = N  \\
                \wedge \, \, a_{2} = ite\left(x[0] > 1, a_{1} - 1, a_{1}\right) \\ 
                \wedge \, \, a_{3} = ite\left(x[1] > 1, a_{2} - 1, a_{2}\right) \\
                \wedge \, \, a_{4} = ite\left(x[2] > 1, a_{3} - 1, a_{3}\right) \\
                \wedge \ldots \\
                \wedge \, \, a_{N+1} = ite\left(x[N-1] > 1, a_{N} - 1 , a_{N}\right) \\
                \wedge \neg \left(a_{N+1} \leq N\right) \\
              \end{array} \right ]  \\
              \\           
\end{equation}

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $a<=N$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can be further transformed into a (equisatisfiable) conjunctive-normal-form formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\mathcal{NP}}$-complete problem~\cite{PatarinG97}).
